\sysfull

### 2017-11-16 paper/systemdescription.tex
overview: shortenning

\sys consists of its compiler and the run time software support for its execution model and includes several lightweight programming model extensions similar to those in prior work~\cite{chain,alpaca}. Table~\ref{tab:viper_syntax} summarizes these extensions, which can be used manually by a programmer or added to a program automatically, using \sys's compiler. As in prior task-based intermittent programming models~\cite{chain,alpaca}, a \sys program is decomposed into tasks, which are functions that are not allowed to have any direct callers. The programmer identifies a task by adding the {\tt task} keyword to the function's declaration. A programmer invokes one task, {\tt t} from another task {\tt u} using {\tt next\_task(\,t\,)}, which atomically commits the state of {\tt u} and transfers control to the entry point of {\tt t}. A program has a single {\em origin} task that executes first during the program's first execution; after a power failure, in subsequent operating periods, \sys restarts from the task that was executing just before the failure.

A \sys program manipulates {\em protected} state---accessed by more than one task---using the {\tt RP} and {\tt WP} operations, which read and write (respectively) a protected variable. These operations convey to \sys's runtime system which data must be protected through paging by the memory manager.

\begin{table}
	\centering
	\footnotesize
	\begin{tabular}{|c|c|}
		\hline
		Keyword & Description\\
		\hline\hline
		\texttt{task foo(){...}} & Task declaration\\
		\texttt{origin task foo(){...}} & Origin task declaration\\
		\texttt{next\_task()} & Transition to a new task\\
		\texttt{RVAR} & Read-protected variable\\
		\texttt{WVAR} & Write-protected variable\\
		\hline
	\end{tabular}
	\caption{\sys's programming model extensions.}
	\label{tab:viper_syntax}
\end{table}

\textbf{\sys Memory Manager.} When the binary executes, \sys's runtime support affects its execution through a {\em (coalescing) task scheduler} (described in Section~\ref{sec:task_coalescing}) and {\em (virtualizing) memory manager} (described in Section~\ref{sec:memory_virtulaization}). The \sys binary executes its tasks correctly, with consistent memory state, due to \sys's virtual memory mechanism. Relying on programmer or compiler annotation of accesses to memory locations that might be shared between tasks, \sys's memory virtualization {\em pages} these data to SRAM from FRAM as they are needed. Tasks operate, accessing SRAM only, demand-paging data to accommodate a SRAM smaller than FRAM. When a task completes, \sys commits modified, paged data back to its original FRAM location atomically ensuring subsequent tasks page in correct memory state.

\textbf{\sys Task Scheduler.} \sys runs a lightweight task scheduler that uses {\em task coalescing} to make \sys's statically defined tasks efficiently use available energy. By default, tasks run in sequence and each task commits as it completes, potentially incurring unnecessary overhead for commits between consecutive, non-failing tasks. \sys dynamically coalesces consecutive tasks by deferring the commit of an earlier task until the completion of a later task. \sys's memory management mechanism facilitates coalescing because \sys can defer an earlier task's commit by simply buffering modified, paged data as the later task executes. After coalescing a sequence of tasks, \sys commits all tasks' updates to FRAM, together, combining tasks updates to the same pages and minimizing the commit overhead.

>>>>

\sys's programming interface, listed in Table~\ref{tab:viper_syntax}, consists of
constructs for defining tasks~\cite{chain,alpaca} and accessors for
protected variables shared between tasks. These anotations are added either by
the programmer (in the manual translation mode) or the compiler (in automatic
translation mode). As in prior task-based intermittent programming
models~\cite{chain,alpaca}, tasks are functions that are not allowed to have
any direct callers but receive control by explicit {\tt next\_task} statements.
In \sys, a task transition atomically commits the memory operations of the
current task.  The programmer identifies each task by adding the {\tt task}
keyword to the function's declaration and designates one task as the {\em
origin} task at which execution begins. After a power failure,
\sys restarts from the last (partially) executed task.
%
A \sys program manipulates {\em protected} state---accessed by more than one
task---using the \emph{read variable} ({\tt RVAR}) and \emph{write variable}
({\tt WVAR}) operations. These operations convey to \sys's memory manager which
data must be buffered in volatile memory and committed.

\begin{table}
	\centering
	\footnotesize
	\begin{tabular}{|c|c|}
		\hline
		\textbf{Keyword} & \textbf{Description} \\
		\hline\hline
		\texttt{task foo() \{...\}} & Task definition\\
		\texttt{origin task foo() \{...\}} & Origin task definition\\
		\texttt{next\_task(t)} & Transition to task \texttt{t}\\
		\texttt{u = RVAR(v)} & Read protected variable \texttt{v} into \texttt{u}\\
		\texttt{WVAR(v, u)} & Write \texttt{u} to protected variable \texttt{v}\\
		\hline
	\end{tabular}
	\caption{\sys's programming interface.}
	\label{tab:viper_syntax}
\end{table}

\textbf{\sys Memory Manager.} At runtime \sys's \emph{virtualizing memory
manager} (Section~\ref{sec:memory_virtulaization}) protects changes to
non-volatile data against becoming inconsistent after a power failure.  Relying
on annotations of accesses to protected variables, \sys's memory manager {\em
pages} these data into volatile memory from non-volatile memory as they are
needed.  Tasks cannot modify protected variables directly in non-volatile
memory.  If the working set exceeds the size of volatile memory, pages are
swapped out to non-volatile memory without committing the changes, leaving
original locations of protected variables unchanged.  When a coalesced task
completes, \sys atomically commits modified pages to their original locations
in non-volatile memory.

\textbf{\sys Task Scheduler.} The scheduler {\em coalesces}
statically\hyp{}defined tasks into dynamic tasks sized to match the available
energy. By default, tasks run in sequence and each task commits as it
completes, potentially incurring unnecessary overhead for commits between
consecutive, non-failing tasks. \sys dynamically coalesces consecutive tasks by
deferring the commit of an earlier task until the completion of a later task.
\sys's memory management mechanism is essential for coalescing, because commits
are deferred by keeping dirty pages in volatile memory across coalesced
boundaries.  After a coalesced sequence of tasks completes, \sys commits pages
updated by all of the tasks to non-volatile memory, reducing the commit
overhead through batching.
